% !TEX program = xelatex

%-------------------------------------%
% Birmingham City University
% Joint Program with University of Information Technology, VNU-HCM
% Module: Individual Honours Project
%
% -----
%
% A Game-Based Approach to Motor Imagery Adaptation and BCI Calibration
% Author: Cao Đăng Khoa (dang.cao@mail.bcu.ac.uk)
% Student ID: 25195675
%
% Supervisor: Dr. Vi Chí Thành (vcthanh@hcmiu.edu.vn)
%------------------------------------%



%------ DOCUMENT CONFIGURATION ------%
\documentclass[a4paper, 11pt, openany]{report}
\usepackage{layout}
\usepackage{fonts}
\usepackage{headerfooter}
\usepackage{frontmatter}
\usepackage{contents}
\usepackage{atbegshi}
\usepackage[round, authoryear]{natbib}
\let\cite\citep

\makeatletter
\let\ps@plain\ps@fancy
\makeatother
%------------------------------------%



%----------- FRONT MATTER -----------%
\begin{document}

% Title Page
\reportTitlePage
    {Project Interim Report}
    {A Game-Based Approach to Motor Imagery Adaptation and BCI Calibration}
    {Cao Đăng Khoa}{25195675}
    {Bachelor of Science with Honours Computer Science}
    {Dr. Vi Chí Thành}{December 1, 2025}
    \thispagestyle{titlepage}

% Page numbering
\pagenumbering{roman}
\setcounter{page}{0}

% Table of Contents
\reportTableOfContents

% List of Figures
\reportListOfFigures

% List of Tables
\reportListOfTables
%------------------------------------%


\cleardoublepage
\pagenumbering{arabic}

\setstretch{1.2} 
\setlength{\parskip}{8pt plus 2pt minus 1pt} 


%------------ CHAPTER 1 -------------%
%----- Introduction and Context -----%
\vspace*{-1.15cm}
\chapter{INTRODUCTION AND CONTEXT}
    % 1.0 Introduction and Context

    \section{Background and Significance}

    The way humans interact with computers has undergone a significant evolution over time. 
    From early keyboards and mice to modern touchscreen devices and voice assistants, 
    each innovation has expanded how people can engage with digital systems. But what 
    if humans were able to communicate with technology simply by thought? This is what 
    next-generation technology is achieving: Brain-Computer Interfaces (BCIs), systems 
    that allow people to control devices with their thoughts.\cite{Birbaumer2006} By 
    reading neural signals directly from the human brain and translating them into 
    digital commands for computers to understand, BCIs not only enhance the experience 
    for typical users but also enable people with physical limitations to access technology 
    more easily.

    Among various types of BCI applications, electroencephalography-based motor imagery 
    (EEG-MI) is one of the most widely studied areas.\cite{Lotte2018} This concept is 
    based on a key principle of neuroscience: whenever a person imagines a movement---
    such as moving a hand or foot---without performing it physically, their brain still 
    produces characteristic signals with patterns that are remarkably similar to the 
    actual physical execution.\cite{Neuper2001} Building on this principle, EEG-based 
    MI has been applied in many real-world scenarios, including motor rehabilitation, 
    assistive control for individuals with disabilities, and control of games.
    \cite{Pfurtscheller2001}

    These promising applications, however, still depend heavily on how accurately MI 
    signals can be detected. This accuracy is influenced by many factors, with the 
    calibration process playing a crucial role.\cite{Lotte2018} Calibration---the 
    supervised training phase where the system learns individual user neural 
    patterns---remains a critical bottleneck limiting practical deployment, which is 
    the main focus of this project.


    % 1.2 Current State and Limitations
    \section{Current State and Limitations}

    In controlled laboratory environments, where the calibration process is managed with precision, EEG-MI systems demonstrate a high level of performance in classifying motor signals.\cite{Lotte2018} After an initial training phase, these classifiers often reach accuracies above 70\% in distinguishing between imagined movements such as left hand, right hand, foot, or a rest state, a level that is commonly taken as the threshold for effective BCI control.\cite{Kübler2001}

    However, achieving this level of performance for each individual user requires a significant calibration effort. In practice, the system must learn a user's specific neural patterns through 15--30 minutes of supervised training, typically collecting 40--80 labelled trials per movement class (around 120--240 trials in total).\cite{Giles2022} Moreover, these EEG patterns vary significantly between sessions even for the same user due to non-stationarities.\cite{Sugiyama2008}

    The repetitive and cognitively demanding nature of EEG-MI makes it challenging to implement outside controlled laboratory settings.\cite{Xu2024} It is time-consuming, requires dedicated human supervision, and depends on specialised lab facilities (often geographically distant). It also induces cognitive fatigue, degrades signal quality over time, and limits practicality for multi-session applications like rehabilitation.\cite{Xu2024} Consequently, despite decades of research, this technology remains largely confined to laboratories rather than achieving widespread real-world adoption.\cite{Wolpaw2002}

    \section{Problem Statement}

    EEG-MI calibration per session requires 15--30 minutes of repetitive trials, which limits real-world scalability despite over 70\% success in the experimental environment.\cite{Giles2022,Kueper2024} This limitation arises from changes in brain signals between sessions and across different individuals,\cite{Sugiyama2008} the need for human supervision,\cite{Giles2022} and reliance on specific infrastructure.\cite{Lotte2018}

%------------------------------------%



%------------ CHAPTER 2 -------------%
%--------- Literature Review --------%
\chapter{Review of Existing Knowledge}
    % 2.0 Review of Existing Knowledge
    Comprehensively outline and expand on the primary concepts and established models presented in
     studies relevant to your problem statement. What is the existing knowledge on this subject? What 
     are the main findings and conclusions derived from these studies? This should not merely be a 
     descriptive summary; rather, a critical analysis of the information to construct a coherent narrative 
     regarding the current state of knowledge. 
    \par 
    
    Note: The review of concepts could be broken down into further sub-sections depending on the themes 
    relevant to your topic that you identify during the review. 

    % 2.1 Critical Analysis and Gap Identification
    \section{Critical Analysis and Gap Identification}
    During this phase, you carry out a critical reflection of the sources you have examined, identifying 
    their limitations, contradictions, or disagreements within the literature. Establish which questions 
    remain unanswered. This critical scrutiny should help you pinpoint a specific "gap”; whether it is an 
    unresolved issue, an underexplored area, a new perspective that has not been considered, or a practical 
    solution that is missing from existing studies.

    % 2.2 Project Justification
    \section{Project Justification}
    Motor imagery-based BCIs are among the most commonly studied paradigms due to their non-invasive nature 
    and potential for rehabilitation applications.
%------------------------------------%



%------------- CHAPTER 3 -------------%
%-Project Aims, Objectives, and Scope-%
\chapter{Project Aims, Objectives, and Scope}
    % 3.0 Project Aims, Objectives, and Scope
    This section clearly defines what your project intends to achieve (aims), the specific steps you will 
    take to reach those goals (objectives), and the boundaries within which your project will operate (scope). 
    \par

    % 3.1 Project Aims
    \section{Project Aims}
    The primary aim of this project is to develop a game-based motor imagery BCI system that can adapt to
        individual users, thereby enhancing calibration efficiency and overall user experience.

    % 3.2 Project Objectives
    \section{Project Objectives}
    To achieve the stated aim, the project will pursue the following objectives:
\section{Project Scope}







%------------ REFERENCES ------------%

\newpage

\chapter*{References}
\renewcommand{\bibname}{\vspace{-1cm}}
\bibliographystyle{apalike}
\bibliography{references}

\end{document}
